Weather Data Integration and Assimilation System\\

### 2020-02-09 method/sec_4-db_struct.tex
Paper - update abstract
Paper - Updated the Structure of the paper

\section{Hierarchical Database Structure}

@@

\subsection{Timeseries Key Attributes}
\label{sub:timeseries_key_attributes}
- Module ID
  String field which describe the source of the data generated. Example of hec-hms, flow2d, weather-station etc
- Value Type
  - Scalar, Vector, Grid are the Value Types which are interested in the WDIAS.
- Location
  Location of the timeseries. All locations have unique String identifier called locationId. Further, there are two type of location types that is interested in the WDIAS.
  - Point locations - Contains a name which is human readable. lat and lon of the location on earth surface.
  \begin{lstlisting}[language=Python]
    {
        "locationId": "wdias-hanwella",
        "name": "Hanwella",
        "lat": 6.909722222,
        "lon": 80.08166667
    }
  \end{lstlisting}
  - Regular Grid locations - Contains a name which is human readable. The grid presents by dividing into equal size cells. Thus, it need the number of rows and columns. 
  And the location of the first cell and the width and height of it.
  \begin{lstlisting}[language=Python]
      {
        "locationId": "wdias_kelani_basin",
        "description": "Kelani Basin",
        "rows": 120,
        "columns": 139,
        "geoDatum": "Kandawala",
        "gridFirstCell": {
            "firstCellCenter": {
                "x": 397074.0,
                "y": 504875.0
            },
            "xCellSize": 250.0,
            "yCellSize": 250.0
        }
    }
\end{lstlisting}
  - Irregular Grid locations (endpoints are available with in the WDIAS system, but skipped since it is out of the interest of the scope.)

@@

\subsection{MySQL}
\label{sub:mysql}
Relational Database Management System (RDBMS).
SQL database has fewer version of open source and it is also possible to gain better performance with commercial versions. 
Since WDIAS scope is to using open source tools, it is using MySQL as persistent database for storing timeseries metadata described above.
One instance of MySQL is using only by adapter-metadata service.
Other than that, another instance of MySQL is using for adapter-extension which will be describe in more detail \ref{data_preprocess}.
MySQL is used only for these services, those will only keep the consistency of the schema for timeseries metadata and extension metadata.
\ref{db_struct_redis} instance is using for caching the data from MySQL, since the MySQL database instances, itself not going to get much load.
Other than MySQL, there are many open source databases can be used to replace the same functionality.

@@

\subsection{MongoDB}
\label{sub:mongodb}
  MongoDB is a general purpose, document-based, distributed database
  Supports query operations on geo-spatial data \cite{mongodbMongoDBManual}
  Clustering and Sharding available for scalability and reliability
MongoDB geo-spatial features are using for support Geo queries in the WDIAS. Other than that, external timeseries metadata queries are served by the indexed timeseries metadata data.
If metadata not present in the adapter-query, if fetch data from the adapter-metadata, then indexed and cache in the MongoDB. Vise a versa, if new timeseries create in the adapter-metadata,
it trigger an event to adapter-query to index the new timeseries.
\ref{micro_y_axis_scaling} concept is using here by splitting into the multiple services based on the functionality and scale the system.

>>>>

\section{Hierarchical Database Structure}
\dbc{Use a more meaningful title.}

@@

\subsection{Timeseries Key Attributes}
\label{sub:timeseries_key_attributes}
- Module ID
  String field which describe the source of the data generated. Example of hec-hms, flow2d, weather-station etc
- Value Type
  - Scalar, Vector, Grid are the Value Types which are interested in the WDIAS.
- Location
  Location of the timeseries. All locations have unique String identifier called locationId. Further, there are two type of location types that is interested in the WDIAS.
  - Point locations - Contains a name which is human readable. lat and lon of the location on earth surface.
  \begin{lstlisting}[language=Python]
    {
        "locationId": "wdias-hanwella",
        "name": "Hanwella",
        "lat": 6.909722222,
        "lon": 80.08166667
    }
  \end{lstlisting}
 \dbc{Line spacing is too high. Use simple monospace format in latex.}

  - Regular Grid locations - Contains a name which is human readable. The grid presents by dividing into equal size cells. Thus, it need the number of rows and columns. 
  And the location of the first cell and the width and height of it.
  \begin{lstlisting}[language=Python]
      {
        "locationId": "wdias_kelani_basin",
        "description": "Kelani Basin",
        "rows": 120,
        "columns": 139,
        "geoDatum": "Kandawala",
        "gridFirstCell": {
            "firstCellCenter": {
                "x": 397074.0,
                "y": 504875.0
            },
            "xCellSize": 250.0,
            "yCellSize": 250.0
        }
    }
\end{lstlisting}
  - Irregular Grid locations (endpoints are available with in the WDIAS system, but skipped since it is out of the interest of the scope.)

@@

\subsection{MySQL}
\label{sub:mysql}

\dbc{Use a term like relational DB tham MySQL}

%Relational Database Management System (RDBMS)
SQL database has fewer version of open source and it is also possible to gain better performance with commercial versions. 
Since WDIAS scope is to using open source tools, it is using MySQL as persistent database for storing timeseries metadata described above.
One instance of MySQL is using only by adapter-metadata service.
Other than that, another instance of MySQL is using for adapter-extension which will be describe in more detail \ref{data_preprocess}.
MySQL is used only for these services, those will only keep the consistency of the schema for timeseries metadata and extension metadata.
\ref{db_struct_redis} instance is using for caching the data from MySQL, since the MySQL database instances, itself not going to get much load.
Other than MySQL, there are many open source databases can be used to replace the same functionality.

@@

\subsection{MongoDB}
\label{sub:mongodb}
\dbc{Use a technology independent term. Same with InfluxDB. MongoDB is used for 2 section names.}

MongoDB is a general purpose, document-based, distributed database
  Supports query operations on geo-spatial data \cite{mongodbMongoDBManual}
  Clustering and Sharding available for scalability and reliability
MongoDB geo-spatial features are using for support Geo queries in the WDIAS. Other than that, external timeseries metadata queries are served by the indexed timeseries metadata data.
If metadata not present in the adapter-query, if fetch data from the adapter-metadata, then indexed and cache in the MongoDB. Vise a versa, if new timeseries create in the adapter-metadata,
it trigger an event to adapter-query to index the new timeseries.
\ref{micro_y_axis_scaling} concept is using here by splitting into the multiple services based on the functionality and scale the system.

### 2020-02-09 method/sec_5-data_preprocess.tex
Paper - Updated the Structure of the paper

\section{Extensions (Weather Data Prepossessing)}
The data imported via imported modules may have incorrect data, missing data and not compatible with directly feed into simulation modules. 
Thus it required to preprocess data before going to use in such cases. The WDIAS supports these capabilities via Extensions. Current system has
in build few extensions such as Interpolation, Transformation and Validation. Other than those, it possible to add more extensions as required.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.5\textwidth]{method/data_preprocess/weather_data_preprocessing.jpg}
    \caption{Weather Data Preprocessing}
    \label{fig:weather_data_preprocessing}
\end{figure}
The WDIAS provides the support via in build simple generic mathematical function structure. Each extension is considering as a mathematical function which can take p 
number of input timeseries variables and output n number of timeseries variables. Other than that, at the time of processing the system can configure to provide
bind constant which can provide while configuring an extension. Which means, it is possible to change the behaviour of a existing function by providing different bind
constant at the time of creating new triggers for an extension.

>>>>

\section{Extensions (Weather Data Prepossessing)}
\dbc{Section title?}

The data imported via imported modules may have incorrect data, missing data and not compatible with directly feed into simulation modules. 
Thus it required to preprocess data before going to use in such cases. The WDIAS supports these capabilities via Extensions. Current system has
in build few extensions such as Interpolation, Transformation and Validation. Other than those, it possible to add more extensions as required.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.5\textwidth]{method/data_preprocess/weather_data_preprocessing.jpg}
    \caption{Weather Data Preprocessing}
    \label{fig:weather_data_preprocessing}
\end{figure}
The WDIAS provides the support via in build simple generic mathematical function structure. Each extension is considering as a mathematical function which can take p 
number of input timeseries variables and output n number of timeseries variables. Other than that, at the time of processing the system can configure to provide
bind constant which can provide while configuring an extension. Which means, it is possible to change the behaviour of a existing function by providing different bind
constant at the time of creating new triggers for an extension.

### 2020-02-09 method/sec_6-query.tex
Before Test plan
Paper - Updated the Structure of the paper

As described in \ref{sub:mongodb}, adapter-query is primarily using for support Geo-spatial queries. Other than that, as the primary service for searching for timeseries, it also supports for retrieval of timeseries metadata for external queries as well.

\subsection{Timeseries Metadata}
Get timeseries metadata by timeseriesID.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries/<TIMESERIES_ID>
\end{lstlisting}
Search for timeseries metadata with key attributes of the timeseries.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries?<QUERY_STRING>
\end{lstlisting}
QUERY\_STRING can be consist of zero to multiple parameters. Multiple query parameters are separated by the ampersand, "\&".
\begin{itemize}
    \item moduleId(Optional) - E.g. HEC-HMS
    \item valueType(Optional) - E.g. Scalar
    \item parameterId(Optional) - E.g. O.Precipitation
    \item locationId(Optional) - E.g. wdias\_hanwella
    \item timeseriesType(Optional) - E.g. ExternalHistorical
    \item timeStepId(Optional) - E.g. each\_15\_min
\end{itemize}
For the moment, query are only support by strict matching with above parameter values.

@@

\subsubsection{Query parameters in Locations}
Using the following API endpoint, it is possible to query the parameters stored against given locations. Location IDs should be provide as list of JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/parameter
    JSON Body:
    ["<locationId1>", ...]
\end{lstlisting}

>>>>

As described in \db{Section} \ref{sub:mongodb}, adapter-query is primarily using for support Geo-spatial queries. Other than that, as the primary service for searching for timeseries, it also supports for retrieval of timeseries metadata for external queries as well.

\subsection{Timeseries Metadata}
Get timeseries metadata by timeseriesID.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries/<TIMESERIES_ID>
\end{lstlisting}
\dbc{Use monospace formatting.}

Search for timeseries metadata with key attributes of the timeseries.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries?<QUERY_STRING>
\end{lstlisting}
QUERY\_STRING can be consist of zero to multiple parameters. Multiple query parameters are separated by the ampersand, "\&".
\begin{itemize}
    \item moduleId(Optional), e.g., HEC-HMS
    \item valueType(Optional), e.g., Scalar
    \item parameterId(Optional), e.g., O.Precipitation
    \item locationId(Optional), e.g., wdias\_hanwella
    \item timeseriesType(Optional), e.g., ExternalHistorical
    \item timeStepId(Optional), e.g., each\_15\_min
\end{itemize}
For the moment, query are only support by strict matching with above parameter values.

@@

\subsubsection{Query parameters in Locations}
Using the following API endpoint, it is possible to query the parameters stored against given locations. Location IDs should be provide as list of JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/parameter
    JSON Body:
    ["<locationID>", ...]
\end{lstlisting}

### 2020-02-09 results/sec_3-observations.tex
Paper - Updated the Structure of the paper

Following \ref{subse:obs_test_plan_all_60min}, \ref{subse:obs_test_plan_all_30min}, \ref{subse:obs_test_plan_all_15min} and \ref{subse:obs_test_plan_all_auto_15min} sections discussed about the observations collected with performing the Full test plan (also known as All test plan) over 60 minutes (24 data points), 30 minutes (48 data points) and 15 minutes (96 data points) by variate the request size. Then another performance test with enabling Auto Scaling as explained in \ref{subse:test_plan_metrics} for high resource utilized microservice.

\subsection{Full Test with hourly data (24 Request Size)}
\label{subse:obs_test_plan_all_60min}
After running the All test plan with 60 minutes (24 data points), it observed the data summary of \ref{tab:obs_all_60_min_summary}. It performed 311k of sample requests.
\begin{table}[]
\begin{tabulary}{\linewidth}{|L|C|C|C|C|C|C|C|C|}
\hline
Label & \# Samples & Avg & Min & Max & 90\% Line & Std. Dev. & Error \% & RPS \\ \hline
Insert Timeseries & 71826 & 28 & 13 & 2773 & 31 & 58.74 & 0.00\% & 40.5 \\ \hline
Retrieve Timeseries & 71796 & 8 & 7 & 242 & 10 & 4.18 & 0.00\% & 40.7 \\ \hline
Insert Grid & 7982 & 23 & 21 & 126 & 26 & 4.23 & 0.06\% & 4.5 \\ \hline
Retrieve Grid & 7979 & 68 & 59 & 238 & 75 & 10.11 & 0.00\% & 4.5 \\ \hline
Query: Location & 71804 & 3 & 2 & 109 & 3 & 1.52 & 0.00\% & 40.5 \\ \hline
\textbf{TOTAL} & 311182 & 127 & 0 & 2773 & 503 & 207.80 & 0.00\% & 175.4 \\ \hline
\end{tabulary}
\caption{Throughput and Latency of All test cases with 60min data}
\label{tab:obs_all_60_min_summary}
\end{table}
Insert Timeseries performed over Scalar and Vector data types with each request send with 24 data points. On average it takes 28ms to insert data into the \acrshort{wdias}, and up to 90\% percentile inserted data within 31ms. Noticeably it has 0\% of errors which means that all the request are completed successfully. On average, \acrshort{wdias} handled 40.5 \acrshort{rps} with the given workload. The performance of the Retrieve of Scalar and Vector Timeseries are similar to insert timeseries due to it depends on it.
On the other hand, Insert Grid Timeseries data perform with 23ms with almost all the request perform within same amount of time, hence the standard deviation is smaller. Retrieve Grid Timeseries data perform the same but with bit higher value of latency of 68ms. Both of insert and retrieve grid timeseries data have 4.5 \acrshort{rps} since it only get 10\% of the number of request per given time via the test plan as explain in the data ration in \ref{se:test_plan}.

>>>>

%Following \ref{subse:obs_test_plan_all_60min}, \ref{subse:obs_test_plan_all_30min}, \ref{subse:obs_test_plan_all_15min} and \ref{subse:obs_test_plan_all_auto_15min} sections
\db{Sections \ref{subse:obs_test_plan_all_60min} to \ref{subse:obs_test_plan_all_auto_15min} discuss} the observations collected with performing the Full test plan (also known as All test plan) over 60 minutes (24 data points), 30 minutes (48 data points) and 15 minutes (96 data points) by variate the request size. Then another performance test with enabling Auto Scaling as explained in \ref{subse:test_plan_metrics} for high resource utilized microservice.

\subsection{Full Test with hourly data (24 Request Size)}
\label{subse:obs_test_plan_all_60min}

\dbc{"Full Test"is a not a good term}

After running the All test plan with 60 minutes (24 data points), it observed the data summary of \ref{tab:obs_all_60_min_summary}. It performed 311k of sample requests.

\dbc{Fonts for tables should be 10pt.}

\begin{table}[]
\begin{tabulary}{\linewidth}{|L|C|C|C|C|C|C|C|C|}
\hline
Label & \# Samples & Avg & Min & Max & 90\% Line & Std. Dev. & Error \% & RPS \\ \hline
Insert Timeseries & 71826 & 28 & 13 & 2773 & 31 & 58.74 & 0.00\% & 40.5 \\ \hline
Retrieve Timeseries & 71796 & 8 & 7 & 242 & 10 & 4.18 & 0.00\% & 40.7 \\ \hline
Insert Grid & 7982 & 23 & 21 & 126 & 26 & 4.23 & 0.06\% & 4.5 \\ \hline
Retrieve Grid & 7979 & 68 & 59 & 238 & 75 & 10.11 & 0.00\% & 4.5 \\ \hline
Query: Location & 71804 & 3 & 2 & 109 & 3 & 1.52 & 0.00\% & 40.5 \\ \hline
\textbf{TOTAL} & 311182 & 127 & 0 & 2773 & 503 & 207.80 & 0.00\% & 175.4 \\ \hline
\end{tabulary}
\caption{Throughput and Latency of All test cases with 60min data}
\label{tab:obs_all_60_min_summary}
\end{table}
Insert Timeseries performed over Scalar and Vector data types with each request send with 24 data points. On average it takes 28ms to insert data into the \acrshort{wdias}, and up to 90\% percentile inserted data within 31ms. Noticeably it has 0\% of errors which means that all the request are completed successfully. On average, \acrshort{wdias} handled 40.5 \acrshort{rps} with the given workload. The performance of the Retrieve of Scalar and Vector Timeseries are similar to insert timeseries due to it depends on it.
On the other hand, Insert Grid Timeseries data perform with 23ms with almost all the request perform within same amount of time, hence the standard deviation is smaller. Retrieve Grid Timeseries data perform the same but with bit higher value of latency of 68ms. Both of insert and retrieve grid timeseries data have 4.5 \acrshort{rps} since it only get 10\% of the number of request per given time via the test plan as explain in the data ration in \ref{se:test_plan}.
